---
title: "Index_heatmap"
author: "Caitie"
date: "2025-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(here)
library(viridis)
library(sf)
```


# Threat index

Themes: climate index
n_ag_A: Agricultural area
n_agc_A: Agricultural census
n_df_A: Deforestation
n_f_Ar: Forestry
n_ls_Ar: Landslides
n_liv_Ar: Livestock
n_mn_Ar: Mining
n_tN_Ar: Nitrogen
n_sed_Ar: Sediment
n_Up_A: Urban population
n_fsh: Fisheries
n_ma_avg_A: Mangroves
Th_A: Average of the above
Tot_A: Average of Th_A and Themes

For index rescaled to priority provinces bordering GSR only

```{r}
threat<-read.csv(here("input_data/TotalIndex_sel1.csv")) %>% 
  filter(ADM2_NA %in% c("Ba", "Ra", "Bua", "Macuata"))


sub_threat<-threat %>%
  dplyr::select(ADM2_NA, ADM3_NA, Ctchm_1, uniq_id2, n_ag_Ar:n_ma_avg_A) 

sub_threat2<-sub_threat %>%
  pivot_longer(cols = n_ag_Ar:n_ma_avg_A, 
               names_to = "Pressure", 
               values_to = "Intensity") %>% 
  mutate(label = ifelse(Pressure == "n_ag_Ar", "Agricultural area (km2)",
                        ifelse(Pressure == "n_agc_A", "Agricultural census",
                        ifelse(Pressure == "n_df_A", "Deforestation",
                        ifelse(Pressure == "n_f_Ar", "Forestry",
                        ifelse(Pressure == "n_ls_Ar", "Landslides",
                        ifelse(Pressure == "n_liv_Ar", "Livestock",
                        ifelse(Pressure == "n_mn_Ar", "Mining",
                        ifelse(Pressure == "n_tN_Ar", "Nitrogen (human wastewater)",
                        ifelse(Pressure == "n_sed_Ar", "Sediment",
                        ifelse(Pressure == "n_Up_A", "Urban population",
                        ifelse(Pressure == "n_fsh", "Fisheries",
                        ifelse(Pressure == "n_ma_avg_A", "Mangrove area (km2)", NA)))))))))))))
```



```{r}

all_plot2<-ggplot(sub_threat2, aes(y=Ctchm_1, x = factor(label, levels = sort(unique(label), decreasing = T)), fill= Intensity)) + 
  geom_tile() +
  theme_bw() +
  scale_fill_viridis(option = "D") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_viridis_c(#trans = "log10",  # Apply log10 transformation to fill
                       option = "D",
                       na.value = "grey50") +     # Choose a viridis option (A, B, C, D))
    labs(x = "Pressure",
        y = "Tikina",
       fill = "Intensity") +
  facet_wrap(~ADM2_NA, scales = "free_y", ncol = 2, axes = "margins", axis.labels = "margins")

all_plot2
```
```{r}
#ggsave(here("figures/Threat_index_heatmap_square.png"), width = 10, height = 10)
```

 # priority basins

```{r}
poi<-read_csv(here("input_data/Top_threat_areas.csv"))
poi_sel<-poi %>%
  filter(bi_class %in% c("3-3", "3-2"))
```

```{r}

sub_threat3<-sub_threat2 %>%
  filter(uniq_id2 %in% poi_sel$uniq_id2)

```

```{r}

all_plot2_poi<-ggplot(sub_threat3, aes(y=Ctchm_1, x = factor(label, levels = sort(unique(label), decreasing = T)), fill= Intensity)) + 
  geom_tile() +
  theme_bw() +
  scale_fill_viridis(option = "D") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_viridis_c(#trans = "log10",  # Apply log10 transformation to fill
                       option = "D",
                       na.value = "grey50") +     # Choose a viridis option (A, B, C, D))
    labs(x = "Pressure",
        y = "Tikina",
       fill = "Intensity") +
  facet_wrap(~ADM2_NA, scales = "free_y", nrow = 3, axes = "margins", axis.labels = "margins")

all_plot2_poi
```
```{r}
ggsave(here("figures/Fig3B_Threat_index_heatmap_square_poi.png"), width = 6, height = 10)
```

```{r}
which(unique(poi$test) %in% unique(sub_threat3$Ctchm_1))
```


# Climate index

```{r}
dat<-read_csv(here("raw_data/Inform_risk_Fiji.csv"))

sub_dat<-dat %>%
  filter(ADM3_NAME %in% c("Ba", "Ra", "Bua", "Macuata")) %>%
  select(-"Total_Population", -"Households",-"THEME1",-"THEME2",-"THEME3", -"THEMES", -"INFORM_risk", -"RISK_class",
         -"Inequality_Index") %>% 
  mutate(Aid_Dependency  = as.numeric(Aid_Dependency))

sub_dat2<-sub_dat %>%
  pivot_longer(cols = Riverine_flood:Lack_of_coping_capacity, 
               names_to = "Factor", 
               values_to = "Vulnerability")
```

```{r}
ggplot(sub_dat2, aes(y=ADM3_NAME, x = factor(Factor, levels = sort(unique(Factor), decreasing = T)), fill= Vulnerability)) + 
  geom_tile() +
  theme_bw() +
  scale_fill_viridis(option = "D") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_viridis_c(trans = "log10",  # Apply log10 transformation to fill
                       option = "D",
                       na.value = "grey50") +     # Choose a viridis option (A, B, C, D))
    labs(x = "Vulnerability variable",
       y = "Tikina",
       fill = "Vulnerability") 
```
```{r}
#ggsave(here("figures/INFORM_heatmap.png"), width = 7, height = 5)
```

